---
title: "Echo_Quality_Dignan"
author: "Lewis Campbell"
date: "3 February 2019"
output: html_document 
    toc: TRUE
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Echo_quality_Dignan
Cahal Dignan's project on the quality control of focused echo in Royal Darwin Hospital ICU

## Background
Echocardiography is used for anatomical, but more often for physiological assessment of the critically ill patient. The echo examination in this setting is generally focused on a clinical question and the most common sonographic sequences focus on size of vessels and cardiac chambers, ventricular contraction and flow patterns across valves or in the great vessels. There are no clearly agreed minimum standards for a focused echo examination. Suggested minimum datasets include that from the College of Intensive Care Medicine of Australia and New Zealand [ref].

Whatever are the sequences acquired, quality assurance is important for clinical safety and the maintenance of professional standards. Images or cine loops must be of adequate quality to detect pathology without the costly and dangerous treatment associated with false positives.

## Setting
RDH ICU is a mixed adult and paediatric ICU, classified as "Metropolitan" in the ANZICS APD to which it is a contributor, but remote from other ICUs and serving a widely dispersed population with high healthcare needs. It is at the time of study a closed unit with 7 FTE Intensivists contributed to by 9 individuals, of whom 4 regularly use ultrasound in practice. Of these 4, one possesses the DDU and two possess other certificates in cardiac sonography (a PGDipEcho and the Glasgow Echo course with a following 10 years of practical experience). Two trainees are in the process of obtaining the CCPU.

Scans are acquired and interpreted by the sonologist, with immediate discussion with another clinician if findings are discordant with the clinical picture or novel and concerning. Later review is limited to reporting at the time of archiving (roughly monthly) and discussion of interesting cases at regular joint echo meetings withe the cardiology department.

## Research questions
1 What volume of scans are performed in RDH ICU under normal conditions?
2 What is the average quality of these scans using a standard measure?
3 What is the influence of sonologist experience and training on that standard measure?

## Methodology 
50 consecutive echo studies were accessed from the archive by an independent researcher (CD) with the prior consent of all sonologists. A standard quality appraisal tool was applied to each echo. Descriptive statistics on adequacy of images were prepared, and the mean values of these per sonologist were calculated. Years of experience and training were summed together for each sonologist, with training arbitrarily taken as twice the value of experience, and the mean quality score was regressed on this "sonologist years" measure.

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
fullresults <- readxl::read_excel("ICU_ECHO.xlsx")
```


## Results
The 50 consecutive scans were performed over 6 months. The infographic displays the time of acquisition of each scan, with a cheeky overlay of quality per sonologist.

```{r}
activity <- ggplot(data=fullresults) +
    geom_histogram(mapping = aes(x=ICU_date, fill = Sonologist, alpha=0.2))
activity
```
There was a suggestion that scan quality was associated with years of sonologist training, with a modest increase of 0.14 on the Quality scale - out of a total score of 13 - per year of training.
```{r, echo=FALSE, warning=FALSE, message=FALSE}

worthit <- lm(Quality ~ Sonologist_years, data=fullresults)
worthit
anova(worthit)
```

The 


## Discussion
Training, even in this small sample, is associated with higher scan Quality. This is generally accepted to be necessary for transparent and reproducible clinical decision making.